---
title: "Final Project"
author: "Line and Cecilie"
date: "2/6/2022"
output: 
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
# load packages:
library(raster) 
library(dplyr) 
library(tidyverse)
library(rworldmap) 
library(sf) 
library(rnaturalearth) 
library(ggplot2) 
library(stats)
library(sp)
library(spData)
# Uncomment this to install package if not already installed
# install.packages("rnaturalearthhires",
#                  repos = "http://packages.ropensci.org",
#                  type = "source")
library(rnaturalearthhires)
#install.packages("ggpubr")
library(ggpubr)

citation("spData")
```



# ANNUAL MEAN TEMPERATURE - WORLD:

```{r world temp }
# Prepare Country Polygon Data
data(world)
plot(world)

# Remove Antarctica
worldClim <- world %>% 
  dplyr::filter(continent!="Antarctica") %>% 
  dplyr::filter(continent!="Seven seas (open ocean)")


# Convert the world object to sp format
global <-as(worldClim,"Spatial")


# Get Climate Data
mean_temp <- getData('worldclim', var='tmean', res=2.5)

# inspect the Climate object
mean_temp

# Convert to Celsius
gain(mean_temp)=0.1
plot(mean_temp)


# Create object that is the annual mean temperature
mean_temp_annual<-mean(mean_temp)

names(mean_temp_annual)<-"mean_temp"


# This next line of code might take a while (approx. 10-15 min). 
# extract values
temp_mean_temp <-raster::extract(x=mean_temp_annual, y=global, fun=mean, na.rm=TRUE, small=TRUE, sp=TRUE)

# Convert the format to sf format
world_temp_sf<-st_as_sf(temp_mean_temp)
```

### Create and save plot of annual mean temperature - world:

```{r plot world temp}

# plot the mean temperature for each country
temp_plot <- ggplot(data=world_temp_sf)+
  geom_sf(aes(fill=mean_temp))+
  scale_fill_viridis_c(name="Annual\nMean\nTemperature (C)", option = "turbo")+
  theme_classic()+
  theme(legend.position="bottom")+
  labs(title="Annual mean temperature - world")+ 
  xlab("Latitude")+
  ylab("Longitude")+
  theme(plot.title = element_text(hjust=0.5))

temp_plot

# Save plot as png
ggsave("temp_plot_world.png", plot = temp_plot, path = "output", width = 12, height = 6, dpi = 300)

```

### Save data in new dataframe:

```{r make df}
# Make into df
mean_temp_countries <- world_temp_sf %>% 
  dplyr::group_by() %>% 
  dplyr::arrange(desc(mean_temp)) %>% 
  dplyr::select(name_long, mean_temp) %>% 
  st_set_geometry(NULL)
```

# ALCOHOL CONSUMPTION - WORLD:

```{r world alc cons}
# CLEAN THE DATA:

# Load alc_cons data
alc_cons <- read_csv("data/alc_cons.csv", show_col_types = FALSE)

# Rename the Consumption column
names(alc_cons)[names(alc_cons) == 'Total alcohol consumption per capita (liters of pure alcohol, projected estimates, 15+ years of age)'] <- 'Consumption'

# Remove rows with NA (We do it this way since there are no other NAs than in the "Code" column)
alc_cons <- na.omit(alc_cons)

# Filter out other years than 2018
alc_cons_clean <- filter(alc_cons, Year == "2018")
```

### Create and save plot of alcohol consumption - world:

```{r plot world alc cons}
# Load world map
world <- ne_countries(scale = "medium", returnclass = "sf") 

# Rename (preparing to merge alcohol data frame with world map data frame)
alc_cons_clean2 <- rename(alc_cons_clean, iso_a3 = Code)

# Merge alcohol data frame with world map data frame
world_alc_merge <- merge(world, alc_cons_clean2, all = TRUE)

# Remove "Antarctica"
world_alc_clean <- world_alc_merge %>% 
  dplyr::filter(continent!="Antarctica")


# Plot global alcohol consumption
cons_map <- ggplot(world_alc_clean)+
  geom_sf(aes(fill=Consumption))+
  scale_fill_viridis_c(name="Alcohol\nconsumption\n(L/capita)", begin = 1, end = 0, option = "magma")+
  theme_classic()+
  theme(legend.position="bottom")+
  labs(title="Alcohol consumption - world")+ 
  theme(plot.title = element_text(hjust=0.5))

cons_map

# Save plot as png
ggsave("alcohol_plot_world.png", plot = cons_map, path = "output", width = 12, height = 6, dpi = 300)

```

# ALCOHOL CONSUMPTION - USA:

```{r usa alc cons}
# ALCOHOL DATA:

# Load US alcohol consumption data

alc_cons_us <- read_csv("data/alc_cons_us.csv", show_col_types = FALSE)

# Convert from gallons to liters
alc_cons_us$alc_cons_L <- alc_cons_us$alcoholConsumptionGallons*3.78541178


# Rename state to entity
names(alc_cons_us)[1] <- 'Entity'




# STATE DATA:

us_states <- ne_states(country = "united states of america", returnclass = "sf")

# rename name column in us_states to "Entity"
names(us_states)[9] <- 'Entity'

# Merge us_states and US alcohol consumption data
us_merge <-merge(us_states, alc_cons_us, by = "Entity")


```

### Create and save plot of alcohol consumption - USA:

```{r  plot usa alc cons}

# Plot US alcohol consumption map
us_cons_map <- ggplot(us_merge)+
  geom_sf(aes(fill=alc_cons_L))+
  coord_sf(xlim=c(-124.7, -67.1), ylim = c(25.2, 49.4), expand = FALSE)+
  scale_fill_viridis_c(name="Alcohol\nconsumption\n(L/capita)", begin = 1, end = 0, option = "magma", 
                       breaks = c(6, 8, 10, 12, 14, 16))+
  theme_classic()+
  theme(legend.position="bottom")+
  labs(title="Alcohol consumption - US")+ 
  theme(plot.title = element_text(hjust=0.5))

us_cons_map

# Save plot as png
ggsave("alcohol_plot_USA.png", plot = us_cons_map, path = "output", width = 12, height = 6, dpi = 300)

```

# ANNUAL MEAN TEMPERATURE: USA:

```{r usa temp}
temp_us <- read_csv("data/temp_us.csv", show_col_types = FALSE)

# Make a column with degrees in celcius
temp_us$temp_c <- (temp_us$AverageTemperature-32)/1.8
# round to 2 decimals
temp_us$temp_c <- round(temp_us$temp_c, 2)

# rename state column in us_states to "Entity"
names(temp_us)[1] <- 'Entity'

#Merge
us_merge <-merge(us_merge, temp_us, by = "Entity")

write.csv(us_merge,"output/us_dataframe.csv", row.names = FALSE)
```


### Create and save plot of anual mean temperature - USA:

```{r plot usa temp}
# Plot US temperature map
us_temp_map <- ggplot(us_merge)+
  geom_sf(aes(fill=temp_c))+
  coord_sf(xlim=c(-124.7, -67.1), ylim = c(25.2, 49.4), expand = FALSE)+
  scale_fill_viridis_c(name="Mean\ntemperature\n(Celcius)", begin = 0, end = 1, option = "turbo")+
  theme_classic()+
  theme(legend.position="bottom")+
  labs(title="Annual mean temperature - US")+ 
  theme(plot.title = element_text(hjust=0.5))

us_temp_map

# Save plot as png
ggsave("temp_plot_USA.png", plot = us_temp_map, path = "output", width = 12, height = 6, dpi = 300)
```


# SPATIAL ANALYSIS (CORRELATION TEST) - WORLD:

```{r merge world data}
# Merge dataframes of Annual Mean Temperature and Alcohol Consumption:

names(mean_temp_countries)[1] <- 'Entity'

data_merge <- merge(mean_temp_countries, alc_cons_clean, by = "Entity")

write.csv(data_merge,"output/world_dataframe.csv", row.names = FALSE)

```

```{r corr test world}
# Scatter plot
cor_test <- ggscatter(data_merge, x = "mean_temp", y = "Consumption", 
          color = "black", shape = 21, size = 3, # Points color, shape and size
          add = "reg.line",  # Add regression line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE, # Add confidence interval
          cor.coef = TRUE, # Add correlation coefficient
          xlab = "Mean temperature (C)", ylab = "Consumption (L/capita)",
          cor.coeff.args = list(method = "spearman", label.x = 22, label.sep = "\n"),
           title = "Correlation test - world")
                      
cor_test

# Save  scatter plot
ggsave("cor_test_world.png", plot = cor_test, path = "output", width = 12, height = 6, dpi = 300)


# Correlation test:
cor.test(x = data_merge$mean_temp, y = data_merge$Consumption, method = "spearman") 
```

# SPATIAL ANALYSIS (CORRELATION TEST) - USA:

```{r corr test usa}
# Scatter plot:
cor_test2 <- ggscatter(us_merge, x = "temp_c", y = "alc_cons_L",
                      color = "black", shape = 21, size = 3, # Points color, shape and size
                      add = "reg.line",  # Add regression line
                      add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                      conf.int = TRUE, # Add confidence interval
                      cor.coef = TRUE, # Add correlation coefficient
                      xlab = "Mean temperature (C)", ylab = "Consumption (L/capita)",
                      cor.coeff.args = list(method = "spearman", label.x =17, label.sep = "\n"),
                      title = "Correlation test - US")

cor_test2

# save scatter plot
ggsave("cor_test_USA.png", plot = cor_test2, path = "output", width = 12, height = 6, dpi = 300)

# Correlation test
cor.test(x = us_merge$temp_c, y = us_merge$alc_cons_L, method = "spearman")
```